---
title: "Introduction to allin1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to allin1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `allin1` provides convenience functions for some common tasks. To install this package from GitHub, use

```{r, eval=FALSE}
# install.packages("remotes")
# install.packages("devtools")
devtools::install_github("ADC-405-S26/allin1")
```

```{r setup}
library(allin1)
```

The package contains three functions

* `sort_mutate()` — creates a new column from an expression and sorts the data
  frame by it in a single step. Sorts ascending by default, or descending with
  `desc = TRUE`. Think of it as `dplyr::mutate()` and `dplyr::arrange()` combined.

* `drop_na_rows()` — removes rows that are *completely* empty (every column is
  `NA`), while keeping rows that have at least one real value. Useful for
  clearing out blank rows without throwing away partially complete records.

* `top_n_percent()` — returns the top *n* percent of rows ranked by a column,
  sorted highest to lowest. For example, the top 50% of employees by hourly
  wage. Passing `percent = 0` returns an empty data frame.
  
## When to use these over base R

Each function wraps a task you *can* already do in base R or dplyr — the point is fewer steps and fewer easy mistakes.

### `sort_mutate()` vs `mutate()` + `arrange()`

Normally you'd compute a column and then sort by it in two separate calls, and it's easy to forget the sort:

```{r, eval=FALSE}
library(dplyr)

employees |>
  mutate(earnings = hours_worked * hourly_wage) |>
  arrange(desc(earnings))
```

`sort_mutate()` does both at once:

```{r, eval=FALSE}
employees |>
  sort_mutate("earnings", hours_worked * hourly_wage, desc = TRUE)
```

### `drop_na_rows()` vs `na.omit()` / `is.na()`

This is the important distinction. `na.omit()` drops a row if it has *any* `NA`. `drop_na_rows()` only drops rows where *every* value is missing, so partially complete rows survive:

```{r}
df <- data.frame(
  name  = c("Alice", "Bob", NA),
  score = c(10,       NA,    NA)
)

# drop_na_rows() keeps Bob, he still has a name
drop_na_rows(df)

# na.omit() removes Bob too, just because his score is missing
na.omit(df)
```

The base-R equivalent of `drop_na_rows()` is `df[rowSums(is.na(df)) < ncol(df), , drop = FALSE]` — correct, but verbose and easy to get wrong (forgetting `drop = FALSE` silently turns a one-column result into a vector).

### `top_n_percent()` vs `slice_max()`

`dplyr::slice_max()` can take a proportion, but `top_n_percent()` lets you think in whole percentages and returns exactly that many rows rather than expanding to include ties:

```{r, eval=FALSE}
# dplyr: proportion as a fraction, and ties may add extra rows
employees |> slice_max(hourly_wage, prop = 0.5)

# allin1: a plain percentage, exact row count
employees |> top_n_percent(hourly_wage, percent = 50)
```

My main motivation comes from everyday problems with data wrangling. People like myself often find it quite tedious and may make silly mistakes while writing large chunks of code like a missing or open pipe operator, or a missing parameter in a function. My goal here was to make shortcuts to more lengthy functions in order to create a more user-friendly data wrangling experience for new and experienced data scientists alike.
