---
title: "Introduction to allin1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to allin1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `allin1` provides convenience functions for some common tasks. To install this package from GitHub, use

```{r, eval=FALSE}
# install.packages("remotes")
# install.packages("devtools")
devtools::install_github("ADC-405-S26/allin1")
```

```{r setup}
library(allin1)
```

The package contains three functions

* `sort_mutate`
* `drop_na_rows`
* `top_n_percent`

My main motivation comes from everyday problems with data wrangling. People like myself often find it quite tedious and may make silly mistakes while writing large chunks of code like a missing or open pipe operator, or a missing parameter in a function. My goal here was to make shortcuts to more lengthy functions in order to create a more user-friendly data wrangling experience for new and experienced data scientists alike.
